---
Learning Base R
Chapter 02
---

Exercise 2.1 ) Write an R command to compute $100*(1 + \dfrac{0.05}{12})^{24}$ which is the value of $100 invest at an annual percentage rate (often abbreviated APR) of 5% after two years with monthly compounding.
```{r}
100 * (1 + 0.05/12)^24
```


Exercise 2.2 ) What is display by the following R command?  (Guess before computing this in R.)
$$ > 5 \%\% 2 $$
```{r}
5 %% 2
```


Exercise 2.3 ) What is the remainder when 3333 is divided by 222?
```{r}
3333 / 222
```


Exercise 2.4 ) Use R to investigate the behavior of $(1 + 1/n)^n$ for large, positive integer values of n
```{r}
(1 + 1/2)^2
(1 + 1/5)^5
(1 + 1/10)^10
(1 + 1/100)^100
```


Exercise 2.5 ) Write R commands to calculate the golden ratio $$\phi = \dfrac{1 + \sqrt{5}}{2}$$ to ten-digit accuracy.
```{r}
options(digits = 10)
(1 + sqrt(5))/2
```


Exercise 2.6 ) The speed of light is 299,792,458 meters per second.  Use R to calculate the amount of enery (the unites will be Joules) in 10 kilograms of mass using Albert Einstein's $$E = m*c^2$$ formula, where m is mass, c is the speed of light and E is the energy.
```{r}
10*299792458^2
```

