---
Learning Base R
Chapter 16
---


Exercise 16.1 ) What is displayed by the following R commands?  (Waking hours from wikipedia.)
```{r}
creatures = c("dog", "cat", "armadillo", "human")
friendly = c(TRUE, TRUE, FALSE, TRUE)
diet = c("cat", "mice", "termites", "Twinkies(tm)")
waking.hours = c(13.9, 11.5, 5.9, 16.0)
creature.data = data.frame(friendly, diet, waking.hours, row.names = creatures)
creatures[creature.data$waking.hours < 12 & ! creature.data$friendly]
```


Exercise 16.2 ) Write a single R command that alphabetizes the rows of the data frame creature.data from Exercise 16.1 by creature name?
```{r}
creature.data[order(rownames(creature.data)), ]
```


Exercise 16.3 ) For the creature.data frame in Exercise 16.1, use the R subset function to create a data frame consisting of just the creature name and diet associated with friendly creatures who are awake more than 12 hours a day.
```{r}
creature.data[creature.data$waking.hours >= 12, ]
```


Exercise 16.4 ) Consider the creature.data data frame in Exercise 16.1
 a.) Write two different R commands that extract the waking hours for a dog.
 b.) Write two different R commands that extract the waking hours for all creatures.
 c.) Write two different R commands that create a data frame that only consists of the rows for dogs and armadillos.
 d.) Write an R command that creates a data frame that consists only of friendly creatures.
```{r}
# a.)
dog <- creature.data["dog", ]
dog$waking.hours

# b.)
waking_hrs <- creature.data$waking.hours
waking_hrs


# c.)
dog_armadillo <- creature.data[rownames(creature.data) %in% c("dog", "armadillo"), ]
dog_armadillo

# d.)
friendly <- creature.data[creature.data$friendly, ]
friendly

```
