---
Machine Learning with R
Chapter 11 - Improving Model Performance
---

  - In this chapter you will learn:
    - How to automate model performance tuning by ssystematically searching for optimal set of training conditions
    - Methods for combining models into groups that use teamwork to tackle tough learning tasks
    - How to apply a variant of decision trees that has quickly become popular due to its impressive performance

## Tuning stock models for better performance

  - <b>parameter tuning</b> : adjusting the model options to identify the best fit.

### Using caret for automated parameter tuning

  - Automated parameter tuning requires you consider three questions:
    - What type of machine learning model (and specific implementation) should be trained on the data?
    - Which model parameters can be adjusted, and how extensively should they be tuned to find the optimal settings?
    - What criteria should be used to evaluate the models to find the best candidate?
  - For a complete list of models and corresponding tuning parameters covered by caret refer to http://topepo.github.io/caret/modelList.html
  - modelLookup() function can be used to lookup parameters for tuning

```{r}
#install.packages("caret")
library(caret)

modelLookup("C5.0")
```

  - By default, caret searches at most three values for each of the models p parameters.
  - The third and final step in automatic model tuning involves identifying the best model among the candidates.

### Creating a simple tuned model

  - We'll use the C5.0 decision trees with the credit model to optimize this model.

```{r}
library(caret)
RNGversion("3.5.2")
set.seed(300)


```

## 

  - 

## 

  - 

## 

  - 

## 

  - 

## 

  - 

## 

  - 

## 

  - 

## 

  - 

### 

```{r}

```

## 

  - 

