---
Machine Learning with R
Chapter 12 - Specialized Machine Learning Topics
---

  - <b>Topics</b>:
    - Unstructured or proprietary formats such as web pages, web APIs, spread-sheets, or databases
    - From a specialized domain such as bioinformatics or social network analysis
    - Too large to fit in memory or unable to complete analyses in a reasonable time

## Managining and preparing real-world data

  - The process of collecting, merging, sorting, filtering and reformatting data to meet requirements is called <b>data mugging</b> or <b>data wrangling</b>.

### Making data "tidy" with the tidyverse packages

  - <b>tidyverse</b>: collection of packages devoted to tidy data.
    - install.packages("tidyverse")

### Generalizing tabular data structures with tibble

  - <b>tibble</b> : acts almost like a data frame put includes additional functionality.
    - Convert a tibble to a data frame, using the as.data.frame() function.
    - Convert a dataframe to a tibble using the as_tibble() function.
  
```{r}
library(tibble)
credit <- read.csv("../ch05/credit.csv")
credit_tbl <- as_tibble(credit)

# - Take a look at the tibble object
credit_tbl
```

### Speeding and simplifying data preparation with dplyr

  - The dplyr package is the heart of tidyverse.
    - Focus's on data frames rather than vectors, new operators are introduced that allow common data transformations.
    - The dplyr package makes reasonable assumptions about data frames
    - Key portions are written in C++.
    - R data frames are limited to available memory.  With dplyr, tibbles can be linked transparently to disk-based databases.
  
  - filter() rows of data by values of columns
  - select() columns of data by name
  - mutate() columns into new columns by transforming the values
  - summarize() rows of data by aggregating values
  - arrange() rows of data by sorting the values
  
  - The 5 dplyr verbs are brought together in  sequences using a <b>pipe</b> operator.  Represented by "%>%"

  - Example : examine loan applicants 21 years or older, and find their average loan duration (in years), grouped by whether or not the applicants defaulted.
```{r}
library(dplyr)

credit_tbl %>%
  filter(age >= 21) %>%
  mutate(years_loan_duration = 
           months_loan_duration / 12) %>%
  select(default, years_loan_duration) %>%
  group_by(default) %>%
  summarize(mean_duration = mean(years_loan_duration))
```

### 

  - 

